### 2024-05-17 2024-5-13/sections/heuristic.tex
add findmodel algorithm

\subsection{Heuristic for RECL Constraints with Nested Counting}\label{heuristic:nested}
 The naive unwinding approach in the paper \cite{Denghang2023} is to unwind all inner counting operators. This approach may generate a large number of states, which is linear to the product of the counting bounds in regex and exponential to the length of the regex. For example, naive unwinding regex $(a^{\{1,1000\}})^{\{1,2\}}$ generate a CEFA containing 1000 states. But if we only unwind the outer counting operator of the regex, not inner, we obtain a CEFA containing only 2 states. The side effect is that we need more registers to store the information of conuting in the inner regex. Recall that unwinding of outer counting operator repeat the inner regex, which also repeat the registers of the inner regex.We present a heuristic to reduce the number of states generated by the unwinding process. The idea is to unwind the counting operator with less bound and less side effect firstly. We present the heuristic in the following algorithm.

@@

The over-approximation is to complement the CEFA of the subregex directly, without unwinding the counting operators in the subregex. The over-approximated complementary algorithm for CEFA is as follows:

@@

\medskip
\noindent
\emph{Step 1. (Compute the transitions times visited by the accepted strings of CEFAs).}

For each CEFA $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$, we already know the values of its registers after solving the existential LIA formula. The values can be seen as a constant vector of length $|R|$. We use $\myvec{R}$ to denote the vector and use $\eta((q, a, q', \myvec{v}))$ to denote the parikh image of $\aut$. Then the number of times that the transition $(q, a, q', \myvec{v})$ visited by the accepted stirng is obtained from the following formula:
$$ \sum \limits_{(q, a, q', \myvec{v})\in \delta} \eta((q, a, q', \myvec{v}))\myvec{v} = \myvec{R}$$

\medskip
\noindent
\emph{Step 2. (Cut off the search tree early by the transitions times visited during the search process).}

After computing the transitions times visited by the accepted strings, we can cut off the search tree of the CEFA. The idea is that we only visit the transition $(q, a, q', \myvec{v})$ if the number of times that the transition visited by the accepted string is greater than $0$. Other transitions are deleted from the CEFA so that we can reduce the search space. Note that the transitions times is updated during the search process so the search space is reduced dynamically. The presudo code of the algorithm is as follows:

>>>>

\subsection{Heuristic for RECL Constraints with Nested Counting}\label{heuristic:nested}
 The naive approach for nested counting in the paper \cite{Denghang2023} is to unwind all inner counting operators. This approach may generate a large number of states, which is linear to the product of the counting bounds in regex and exponential to the length of the regex. For example, naive unwinding regex $(a^{\{1,1000\}})^{\{1,2\}}$ generate a CEFA containing 1000 states. But if we only unwind the outer counting operator of the regex, not inner, we obtain a CEFA containing only 2 states. The side effect is that we need two registers to store the information of conuting in the inner regex (unwinding of outer counting operator repeat the inner regex, which also repeat the registers of the inner regex).We present a heuristic to reduce the number of states generated by the unwinding process. The idea is to unwind the counting operator with less bound and less side effect firstly. We present the heuristic in the following algorithm.

@@

The over-approximation is to "complement" the CEFA of the subregex directly, without unwinding the counting operators in the subregex. The over-approximated complementary algorithm for CEFA is as follows:

@@

\medskip
\noindent
\emph{Step 1. (Compute the transitions times visited by the accepted string of CEFAs).}

For each CEFA $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$, we already know the values of its registers after solving the existential LIA formula. The values can be seen as a constant vector of length $|R|$. We use $\myvec{R}$ to denote the vector and use $\eta((q, a, q', \myvec{v}))$ to denote the parikh image of $\aut$. Then the transitions times visited by the accepted stirng is obtained from the following formula:
$$ \sum \limits_{(q, a, q', \myvec{v})\in \delta} \eta((q, a, q', \myvec{v}))\myvec{v} = \myvec{R}$$

\medskip
\noindent
\emph{Step 2. (Prune the search tree early by the transitions times during the search process).}

After computing the transitions times visited by the accepted string, we prune the search tree of the CEFA based on the transitions times. The idea is that we only visit transitions whose visited times are greater than $0$. Other transitions are deleted from the CEFA so that we can reduce the search space. Note that the transitions times is updated during the search process so the search space is reduced dynamically. The presudo code is presented in Algorithm \ref{alg:findword}.

\begin{algorithm}
  \caption{Find a word in the language of a CEFA}\label{alg:findword}
  \begin{algorithmic}[1]
  \Procedure{findmodel}{$\aut,tmap$}\Comment{$tmap$ maps transition to its visited times}
      \State $todo \gets Stack(\aut_{init}, tmap, '''')$\Comment{todo is a stack of 3-tuple (state, tmap, word)}
      \State $visited \gets \emptyset$\Comment{visited is a set of 2-tuple (state, tmap)}
      \While{$todo$ is not empty}
        \State $(q, tmap, w) \gets todo.pop()$
        \If{$q\in \aut_{final}$ and all times of transitions in $tmap$ are $0$}
          \State \Return $w$
        \EndIf
        \For{$(q, a, q', \myvec{v})\in \delta$}
          \State $tmap'[(q, a, q', \myvec{v})] \gets tmap[(q, a, q', \myvec{v})] - 1$
          \If{$tmap'[(q, a, q', \myvec{v})] < 0$}
            \State \textbf{continue}\Comment{Prune}
          \EndIf
          \If{$(q', tmap')\in visited$}
            \State \textbf{continue}\Comment{Prune}
          \EndIf
          \If{there exisits $t$ with $tmap[t]>0$ not reachable from $q'$}
            \State \textbf{continue}\Comment{Prune}
          \EndIf
          \State $todo.push((q', tmap', w+a))$
          \State $visited.add((q', tmap'))$
        \EndFor
      \EndWhile
      \State \Return $\emptyset$ 
  \EndProcedure
  \end{algorithmic}
  \end{algorithm}

  The procedure \textsc{findmodel} takes a CEFA $\aut$ and a map $tmap$ as input. The map $tmap$ maps each transition to its visited times. Stack $todo$ is used to store the state need to be visited and its corresponding map and word. The procedure uses a set $visited$ to store the state and its corresponding map that have been visited. In the while loop from line 4 to line 23, the procedure depth-first search the CEFA. If the state is an accepting state and all transitions are visited with correct times, the procedure returns the word (line 7). During the search process, the procedure prunes the search space by the transitions times. If the visited times of a transition is less than $0$, the procedure does not visit it (line 11). If the state and its corresponding map have been visited, the procedure does not visit it (line 14). If there is a transition with visited times greater than $0$ not reachable from the current state, the procedure does not visit it(line 18). In practice, the pruning in line 18 make a great contribution to the efficiency of the search process.
